---
title: "Model selection 10/2020"
output: github_document
---

Including history of the crossval approach, how I've arrived at a combined crossval + AIC approach, and what I see as paths from here....


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library(LDATS)
source(here::here("analysis", "fxns", "crossval_fxns.R"))
source(here::here("analysis", "fxns", "hybrid_fxns.R"))

load(here::here("old_but_mb_useful", "bbs_1_11.RData"))
```

As an example, working with this BBS dataset (`rtrg_1_11`):

```{r plot bbs data}

long_abund <- bbs_rtrg_1_11$abundance %>%
  mutate(year = bbs_rtrg_1_11$covariates$year) %>%
  tidyr::pivot_longer(-year, names_to = "species", values_to ="abundance") %>%
  group_by(year) %>%
  mutate(total_abund = sum(abundance)) %>%
  ungroup()%>%
  mutate(prop_abund = abundance/total_abund)

ggplot(long_abund, aes(year, prop_abund, color = species)) +
  geom_line() +
  theme_bw() +
  scale_color_viridis_d(end = .9) +
  theme(legend.position = "none")

```
### Classic LDATS

Initially, LDATs was set up to select a combination of an LDA and a TS model sequentially using AIC(c). First we fit many LDA models with a variety of `k` and `seed`, and select the best-fitting LDA model based on AIC. Then we use the topic proportions (`gamma`) from that model to fit several TS models with a variety of `nchangepoints` and `formula`. We select the best TS model based on AIC(c). 

```{r classic LDATs LDAs, echo = T}

classic_LDAs <- LDATS::LDA_set(bbs_rtrg_1_11$abundance, topics = c(2,3,5,6,9,11,13,15), nseeds = 1)

classic_LDA_select <- LDATS::select_LDA(classic_LDAs)

plot_lda_comp(classic_LDA_select)
plot_lda_year(classic_LDA_select, bbs_rtrg_1_11$covariates$year)
```

Right away this illustrates an issue, which we did not have with Portal. The best-fitting LDA, using AIC, has **eleven** topics. This is dimensionality reduction from >80 species, which is what we start with in this dataset, but is still not intuitive. Also, as is evident from the plot of the gammas over time, the topics behave idiosyncratically over time and several topics appear to be precisely fit to one or a few time steps. These are **very complicated** temporal dynamics, and it's going to be hard for a changepoint model to distill any kind of pattern out of them:

```{r classic LDATs TS}

classic_TS <- LDATS::TS_on_LDA(classic_LDA_select, as.data.frame(bbs_rtrg_1_11$covariates), formulas = ~ 1, nchangepoints = c(0:3), timename = "year", control = LDATS::TS_control(nit = 100))

classic_TS_select <- LDATS::select_TS(classic_TS)

gamma_plot(classic_TS_select)

```
