---
title: "Hybrid results 10/2020"
output: github_document
---

Including history of the crossval approach, how I've arrived at a combined crossval + AIC approach, and what I see as paths from here....


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(ggplot2)
library(LDATS)
source(here::here("analysis", "fxns", "crossval_fxns.R"))
source(here::here("analysis", "fxns", "hybrid_fxns.R"))
source(here::here("analysis", "fxns", "make_toy_data_objects.R"))
```

```{r}

all_evals <- read.csv(here::here("analysis", "all_evals_hybrid.csv"))

all_evals <- all_evals %>%
  mutate(k_seed = paste(k,seed, sep = "_")) %>%
  filter(grepl(dataset, pattern = "toy_new"))
  

ggplot(all_evals, aes(as.factor(k), sum_loglik, color = as.factor(cpts), group = k_seed)) + geom_boxplot() + facet_wrap(vars(dataset), scales = "free_y")

# there is a lot of spread per model, and overlap.
# taking this as NOT problematic....precedent, AIC is often calculated off the mean.

# look at the mean

all_evals_summary <- all_evals %>%
  group_by(dataset, k, seed, cpts) %>%
  summarize(mean_sum_ll = mean(sum_loglik)) %>%
  arrange(desc(mean_sum_ll)) %>%
  group_by(dataset) %>%
  mutate(dat_rank = row_number())


ggplot(filter(all_evals_summary), aes(as.factor(k), mean_sum_ll, color = as.factor(cpts))) + 
  geom_point() + 
  facet_wrap(vars(dataset), scales = "free_y")

# Some but not all have clear winners. 
(filter(all_evals_summary, dat_rank < 6))

```

## New static
```{r static}

static <- get_toy_data("new_static", here::here("analysis", "toy_datasets"))

static_long <- static$abundance %>%
  mutate(year = static$covariates$year) %>%
  tidyr::pivot_longer(-year, names_to = "species", values_to = "abundance")

ggplot(static_long, aes(year, abundance, color = species)) +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_viridis_d(end = .8)

(filter(all_evals_summary, dat_rank < 6, dataset == "all_evals_toy_new_static"))

static_ldas <- LDATS::LDA_set_user_seeds(static$abundance,  topics = c(3, 5, 8, 9), seed = 16)

for(i in 1:4) {
  print(plot_lda_year(static_ldas[[i]], static$covariates$year))
}

static_ts <- LDATS::TS_on_LDA(static_ldas, as.data.frame(static$covariates), formulas = ~1, nchangepoints = 0,timename = "year", control = LDATS::TS_control(nit = 100))

for(i in 1:4) {
  print(gamma_plot(static_ts[[i]], static$covariates$year))
}
```

## New changepoint
```{r changepoint}

changepoint <- get_toy_data("new_changepoint", here::here("analysis", "toy_datasets"))

changepoint_long <- changepoint$abundance %>%
    mutate(year = changepoint$covariates$year) %>%
  tidyr::pivot_longer(-year, names_to = "species", values_to = "abundance")

ggplot(changepoint_long, aes(year, abundance, color = species)) +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_viridis_d(end = .8)

(filter(all_evals_summary, dat_rank < 6, dataset == "all_evals_toy_new_changepoint"))


changepoint_ldas <- list(LDATS::LDA_set_user_seeds(changepoint$abundance,  topics = 5, seed = 18),
                    LDATS::LDA_set_user_seeds(changepoint$abundance,  topics = 2, seed = 20))


for(i in 1:length(changepoint_ldas)) {
  print(plot_lda_year(changepoint_ldas[[i]], changepoint$covariates$year))
}

changepoint_ts_5 <- LDATS::TS_on_LDA(changepoint_ldas[[1]], as.data.frame(changepoint$covariates), formulas = ~1, nchangepoints = 1,timename = "year", control = LDATS::TS_control(nit = 100))
                       
changepoint_ts_2 <- LDATS::TS_on_LDA(changepoint_ldas[[2]], as.data.frame(changepoint$covariates), formulas = ~1, nchangepoints = 1,timename = "year", control = LDATS::TS_control(nit = 100))

gamma_plot(changepoint_ts_5[[1]])
rho_plot(changepoint_ts_5[[1]]) +
  xlim(0, nrow(changepoint$abundance))
gamma_plot(changepoint_ts_2[[1]])
rho_plot(changepoint_ts_2[[1]]) +
  xlim(0, nrow(changepoint$abundance))

```


## New two changepoint
```{r two changepoint}

two_changepoint <- get_toy_data("new_two_changepoint", here::here("analysis", "toy_datasets"))

two_changepoint_long <- two_changepoint$abundance %>%
    mutate(year = two_changepoint$covariates$year) %>%
  tidyr::pivot_longer(-year, names_to = "species", values_to = "abundance")

ggplot(two_changepoint_long, aes(year, abundance, color = species)) +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_viridis_d(end = .8)


ggplot(two_changepoint_long, aes(year, abundance, color = species)) +
  geom_line() +
  theme_bw() +
  theme(legend.position = "none") +
  scale_color_viridis_d(end = .8) +
  facet_wrap(vars(species), scales = "free_y")

(filter(all_evals_summary, dat_rank < 6, dataset == "all_evals_toy_new_two_changepoint"))


two_changepoint_ldas <- list(LDATS::LDA_set_user_seeds(two_changepoint$abundance,  topics = 2, seed = 12),
                    LDATS::LDA_set_user_seeds(two_changepoint$abundance,  topics = 3, seed = 4),
                    LDATS::LDA_set_user_seeds(two_changepoint$abundance,  topics = 3, seed = 12))


for(i in 1:length(two_changepoint_ldas)) {
  print(plot_lda_year(two_changepoint_ldas[[i]], two_changepoint$covariates$year))
}

two_changepoint_ts_2 <- LDATS::TS_on_LDA(two_changepoint_ldas[[1]], as.data.frame(two_changepoint$covariates), formulas = ~1, nchangepoints = 1,timename = "year", control = LDATS::TS_control(nit = 100))
                       
two_changepoint_ts_3 <- LDATS::TS_on_LDA(two_changepoint_ldas[[2]], as.data.frame(two_changepoint$covariates), formulas = ~1, nchangepoints = 1,timename = "year", control = LDATS::TS_control(nit = 100))

two_changepoint_ts_3_2cpts <- LDATS::TS_on_LDA(two_changepoint_ldas[[3]], as.data.frame(two_changepoint$covariates), formulas = ~1, nchangepoints = 2,timename = "year", control = LDATS::TS_control(nit = 100))

gamma_plot(two_changepoint_ts_2[[1]])
rho_plot(two_changepoint_ts_2[[1]]) +
  xlim(0, nrow(two_changepoint$abundance))
gamma_plot(two_changepoint_ts_3[[1]])
rho_plot(two_changepoint_ts_3[[1]]) +
  xlim(0, nrow(two_changepoint$abundance))
gamma_plot(two_changepoint_ts_3_2cpts[[1]])
rho_plot(two_changepoint_ts_3_2cpts[[1]]) +
  xlim(0, nrow(two_changepoint$abundance))
```
  
For the data I (intended) to contrive to have two changepoints, the LDA seems to have found a way to split into just **2** topics that capture most of the change. I am not sure how it has done this, but I'm not completely surprised. It is very hard to engineer topic structure over time. 
